---
title: "Crime Data"
author: "Alex Hentz"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
crime <- read.csv("C:/Users/alexh/Downloads/Crime_Data_from_2020_to_Present.csv")
View(crime)
crime_2 <- read.csv("C:/Users/alexh/Downloads/Crime_Data_from_2010_to_2019.csv")
View(crime_2)
library(tidyverse)
library(dplyr)

# Top 15 types of Crime in the City of Los Angeles

N <- 15

crime <- mutate(crime, Crm.Cd.Desc = as.character(Crm.Cd.Desc))

common_crime_dest <- crime %>% count(Crm.Cd.Desc, sort = TRUE)

most_common_dest <- head(common_crime_dest, N)

cat("Top", N, "most common Crm.Cd.Desc are:\n")
print(most_common_dest[, c("Crm.Cd.Desc", "n")])


# Plot the top 15 types of Crime reported 
library(ggplot2)
library(plotly)

plot_ly(data = most_common_dest, x = ~Crm.Cd.Desc, y = ~n, type = "bar", marker = list(color = "skyblue")) %>%
  layout(title = paste("Top", N, "Most Common Crm.Cd.Desc"),
         xaxis = list(title = "Crime Description"),
         yaxis = list(title = "Count"))


# Split them up by year and the month
library(lubridate)

N <- 15

crime <- mutate(crime, Date.Rptd = as.character(Date.Rptd))

crime <- mutate(crime, ParsedDate = mdy_hms(Date.Rptd))

crime <- mutate(crime, Month = month(ParsedDate), Year = year(ParsedDate))

crime <- mutate(crime, Crm.Cd.Desc = as.character(Crm.Cd.Desc))

common_crime_dest <- crime %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

top_crimes_by_month_year <- common_crime_dest %>%
  group_by(Year, Month) %>%
  slice_head(n = N)

print(top_crimes_by_month_year[, c("Year", "Month", "Crm.Cd.Desc", "n")])

# Combine the months of each year
N <- 15

crime <- mutate(crime, Date.Rptd = as.character(Date.Rptd), Crm.Cd.Desc = as.character(Crm.Cd.Desc))

crime <- mutate(crime, ParsedDate = lubridate::mdy_hms(Date.Rptd))

crime <- mutate(crime, Month = month(ParsedDate), Year = year(ParsedDate))

common_crime_dest <- crime %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

top_crimes_by_month_year <- common_crime_dest %>%
  group_by(Year, Crm.Cd.Desc) %>%
  summarize(TotalCount = sum(n)) %>%
  arrange(desc(TotalCount)) %>%
  slice_head(n = N)

print(top_crimes_by_month_year)

# Plot the Results on a Bar Chart
plot_2 <- plot_ly(top_crimes_by_month_year, x = ~as.factor(Year), y = ~TotalCount, color = ~Crm.Cd.Desc, type = 'bar') %>%
  layout(barmode = 'stack', title = 'Top 15 Crimes by Year')

plot_2

# Cleaning Crime_2 data
N <- 15

crime_2 <- mutate(crime_2, Crm.Cd.Desc = as.character(Crm.Cd.Desc))

common_crime_dest_2 <- crime_2 %>% count(Crm.Cd.Desc, sort = TRUE)

most_common_dest_2 <- head(common_crime_dest_2, N)

cat("Top", N, "most common Crm.Cd.Desc are:\n")
print(most_common_dest_2[, c("Crm.Cd.Desc", "n")])



N <- 15

crime_2 <- mutate(crime_2, Date.Rptd = as.character(Date.Rptd))

crime_2 <- mutate(crime_2, ParsedDate = mdy_hms(Date.Rptd))

crime_2 <- mutate(crime_2, Month = month(ParsedDate), Year = year(ParsedDate))

crime_2 <- mutate(crime_2, Crm.Cd.Desc = as.character(Crm.Cd.Desc))

common_crime_dest_2 <- crime_2 %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

top_crimes_by_month_year_2 <- common_crime_dest_2 %>%
  group_by(Year, Month) %>%
  slice_head(n = N)

print(top_crimes_by_month_year_2[, c("Year", "Month", "Crm.Cd.Desc", "n")])

N <- 15

crime_2 <- mutate(crime_2, Date.Rptd = as.character(Date.Rptd), Crm.Cd.Desc = as.character(Crm.Cd.Desc))

crime_2 <- mutate(crime_2, ParsedDate = lubridate::mdy_hms(Date.Rptd))

crime_2 <- mutate(crime_2, Month = month(ParsedDate), Year = year(ParsedDate))

common_crime_dest_2 <- crime_2 %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

top_crimes_by_month_year_2 <- common_crime_dest_2 %>%
  group_by(Year, Crm.Cd.Desc) %>%
  summarize(TotalCount = sum(n)) %>%
  arrange(desc(TotalCount)) %>%
  slice_head(n = N)

print(top_crimes_by_month_year_2)

# Plot for Crime_2 
plot_4 <- plot_ly(top_crimes_by_month_year_2, x = ~as.factor(Year), y = ~TotalCount, color = ~Crm.Cd.Desc, type = 'bar') %>%
  layout(barmode = 'stack', title = 'Top 15 Crimes by Year')

plot_4

# Merge Crime 1 and 2 data sets
all_crime <- rbind(crime, crime_2)

# Show top 15 from 2010 to Present
N <- 30

all_crime <- mutate(all_crime, Crm.Cd.Desc = as.character(Crm.Cd.Desc))

common_crime_dest_3 <- all_crime %>% count(Crm.Cd.Desc, sort = TRUE)

most_common_dest_3 <- head(common_crime_dest_3, N)

N <- 30

all_crime <- mutate(all_crime, Date.Rptd = as.character(Date.Rptd))

all_crime <- mutate(all_crime, ParsedDate = mdy_hms(Date.Rptd))

all_crime <- mutate(all_crime, Month = month(ParsedDate), Year = year(ParsedDate))

all_crime <- mutate(all_crime, Crm.Cd.Desc = as.character(Crm.Cd.Desc))

common_crime_dest_3 <- all_crime %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

top_crimes_by_month_year_3 <- common_crime_dest_3 %>%
  group_by(Year, Month) %>%
  slice_head(n = N)

N <- 30

all_crime <- mutate(all_crime, Date.Rptd = as.character(Date.Rptd), Crm.Cd.Desc = as.character(Crm.Cd.Desc))

all_crime <- mutate(all_crime, ParsedDate = lubridate::mdy_hms(Date.Rptd))

all_crime <- mutate(all_crime, Month = month(ParsedDate), Year = year(ParsedDate))

common_crime_dest_3 <- all_crime %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

top_crimes_by_month_year_3 <- common_crime_dest_3 %>%
  group_by(Year, Crm.Cd.Desc) %>%
  summarize(TotalCount = sum(n)) %>%
  arrange(desc(TotalCount)) %>%
  slice_head(n = N)


N <- 30

all_crime <- mutate(all_crime, Date.Rptd = as.character(Date.Rptd), Crm.Cd.Desc = as.character(Crm.Cd.Desc))

all_crime <- mutate(all_crime, ParsedDate = lubridate::mdy_hms(Date.Rptd))

all_crime <- mutate(all_crime, Month = month(ParsedDate), Year = year(ParsedDate))

common_crime_dest_3 <- all_crime %>% count(Year, Month, Crm.Cd.Desc, sort = TRUE)

# Plot 
plot_3 <- plot_ly(top_crimes_by_month_year_3, x = ~as.factor(Year), y = ~TotalCount, color = ~Crm.Cd.Desc, type = 'bar') %>%
  layout(barmode = 'stack', title = 'Top 15 Crimes by Year')

plot_3